\section{Very Short Overview}

\begin{enumerate}
    \item Consideration:
    \begin{enumerate}
        \item Donative promises, gifts, conditional donative promises.
        \item Form.
        \item Reliance, estoppel in pais, promissory estoppel.
        \item The bargain principle, adequacy, duress, salvage.
        \item Unconscionability.
        \item Mutuality, illusory promises.
        \item Legal duty, extortion.
        \item Modification, waiver.
        \item Past consideration, moral obligations.
    \end{enumerate}
    \item Remedies:
    \begin{enumerate}
        \item Efficient breach.
        \item Expectation---services, sale of goods (buyer's and seller's 
        breach), mitigation, foreseeability, certainty, liquidated damages.
        \item Specific performance, uncertainty, uniqueness.
        \item Reliance.
        \item Restitution, unjust enrichment, quantum meruit.
    \end{enumerate}
    \item Assent:
    \begin{enumerate}
        \item Offer and revocation:
        \begin{enumerate}
            \item Interpretation---objectivity vs. subjectivity, meeting of 
            the minds, reasonableness, trade usage.
            \item Offer---Negotiation, firm offer, unilateral offer.
            \item Lapse, rejection, qualified acceptance, renewal.
            \item Revocation---unilateral contract rule, 
            consideration/reliance, subcontractors, option contracts.
        \end{enumerate}
        \item Acceptance by act, subjective acceptance, acceptance by conduct, 
        using a subcontractor's bid, silence, electronic agent.
        \item Implied-in-law, implied-in-fact.
        \item Preliminary negotiations, indefiniteness, gap-fillers, future 
        agreements, good faith negotiations.
        \item Parol evidence, lawyer's paradise, Williston vs. Corbin, 
        merger/integration clauses, fraud exception, 
        condition-to-legal-effectiveness exception.
        \item Plain meaning, Traynor vs. Kozinski.
    \end{enumerate}
    \item Form contracts:
    \begin{enumerate}
        \item Battle of the forms, UCC \S\ 2-207 (``expressly conditional''), 
        knockout rule.
        \item Rolling contracts, shrinkwrap agreements, clickwrap agreements.
        \item Reformation, unconscionability.
        \item Llewellyn.
    \end{enumerate}
    \item Mistake and unexpected circumstances:
    \begin{enumerate}
        \item Unilateral mistake/mechanical error, mistake in transcription.
        \item Mutual mistake, impossibility, \emph{Sherwood}, ``basic 
        assumption.''
        \item Unexpected circumstances, implied conditions, tacit assumptions, 
        impracticability.
        \item Implied covenant of good faith and fair dealing, difficulty of 
        performance.
    \end{enumerate}
\end{enumerate}
